Spatiotemporal Optimization of Cosmological Particel-Mesh Simulation

### 2022-12-01 docs/papers/sto/sto.tex
FIXUP draft: add some obj text

%FIXME copied from ../adjoint/adjoint.tex
\newcommand{\deltaD}{\delta^\textsc{d}}
\newcommand{\deltaK}{\delta^\textsc{k}}
\renewcommand{\d}{d}
\newcommand{\p}{\partial}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\cR}{\mathcal{R}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\cH}{\mathcal{H}}
\bmdefine{\vzero}{0}
\bmdefine{\vI}{I}
\bmdefine{\vnabla}{\nabla}
\bmdefine{\vtheta}{\theta}  % parameters
\bmdefine{\vomega}{\omega}  % white noise modes
\bmdefine{\vk}{k}  % wavevectors
\bmdefine{\vx}{x}  % comoving and canonical coordinates
\bmdefine{\vq}{q}  % Lagrangian coordinates
\bmdefine{\vs}{s}  % displacements
\bmdefine{\vp}{p}  % canonical momenta
\bmdefine{\va}{a}  % accelerations
\bmdefine{\vz}{z}  % states
\bmdefine{\vf}{f}
\bmdefine{\vF}{F}
\bmdefine{\vDelta}{\Delta}
\bmdefine{\vLambda}{\Lambda}
\bmdefine{\vlambda}{\lambda}
\bmdefine{\vvarphi}{\varphi}
\bmdefine{\vxi}{\xi}  % x adjoint
\bmdefine{\vpi}{\pi}  % p adjoint
\bmdefine{\valpha}{\alpha}  % force vjp x gradient
\bmdefine{\vzeta}{\zeta}  % force vjp theta gradient
\newcommand{\half}{\nicefrac12}
\newcommand{\As}{A_\mathrm{s}}
\newcommand{\ns}{n_\mathrm{s}}
\newcommand{\Omegam}{\Omega_\mathrm{m}}
\newcommand{\Omegab}{\Omega_\mathrm{b}}
\newcommand{\Mpc}{\mathrm{Mpc}}
\newcommand{\ic}{\mathrm{i}}
\newcommand{\Plin}{P_\mathrm{lin}}

@@

Given two snapshots at $a_i$ and $a_{i+1}$, we get the particle displacements
in between at $a$ using interpolation with the cubic Hermite spline,
\begin{align}
  \bm{x}(a) =\ &h_{00}(\alpha)\bm{x}_i + h_{10}(\alpha)(a_{i+1} - a_i)\bm{v}_i + \nonumber\\
           &h_{01}(\alpha)\bm{x}_{i+1} + h_{11}(\alpha)(a_{i+1} - a_i)\bm{v}_{i+1} \,,
\end{align}
where $\alpha := (a - a_i)/(a_{i+1} - a_i)$ and the Hermite basis functions read
\begin{align}
  h_{00}(\alpha) &= 2\alpha^3 - 3\alpha^2 + 1 \,,\nonumber\\
  h_{10}(\alpha) &= \alpha^3 - 2\alpha^2 + \alpha \,,\nonumber\\
  h_{01}(\alpha) &= -2\alpha^3 + 3\alpha^2 \,,\nonumber\\
  h_{11}(\alpha) &= \alpha^3 - \alpha^2 \,.
\end{align}
The particle velocity is then given by the derivative $d\bm{x}/da$.

>>>>

%FIXME copied from ../adjoint/adjoint.tex
\newcommand{\deltaD}{\delta^\textsc{d}}
\newcommand{\deltaK}{\delta^\textsc{k}}
\renewcommand{\d}{d}
\newcommand{\p}{\partial}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\cR}{\mathcal{R}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\cH}{\mathcal{H}}
\bmdefine{\vzero}{0}
\bmdefine{\vI}{I}
\bmdefine{\vnabla}{\nabla}
\bmdefine{\vtheta}{\theta}  % parameters
\bmdefine{\vomega}{\omega}  % white noise modes
\bmdefine{\vk}{k}  % wavevectors
\bmdefine{\vx}{x}  % comoving and canonical coordinates
\bmdefine{\vq}{q}  % Lagrangian coordinates
\bmdefine{\vs}{s}  % displacements
\bmdefine{\vp}{p}  % canonical momenta
\bmdefine{\vv}{v}  % FIXME some velocities
\bmdefine{\va}{a}  % accelerations
\bmdefine{\vz}{z}  % states
\bmdefine{\vf}{f}
\bmdefine{\vF}{F}
\bmdefine{\vDelta}{\Delta}
\bmdefine{\vLambda}{\Lambda}
\bmdefine{\vlambda}{\lambda}
\bmdefine{\vvarphi}{\varphi}
\bmdefine{\vxi}{\xi}  % x adjoint
\bmdefine{\vpi}{\pi}  % p adjoint
\bmdefine{\valpha}{\alpha}  % force vjp x gradient
\bmdefine{\vzeta}{\zeta}  % force vjp theta gradient
\newcommand{\half}{\nicefrac12}
\newcommand{\As}{A_\mathrm{s}}
\newcommand{\ns}{n_\mathrm{s}}
\newcommand{\Omegam}{\Omega_\mathrm{m}}
\newcommand{\Omegab}{\Omega_\mathrm{b}}
\newcommand{\Mpc}{\mathrm{Mpc}}
\newcommand{\ic}{\mathrm{i}}
\newcommand{\Plin}{P_\mathrm{lin}}

@@

Given two snapshots at $a_i$ and $a_{i+1}$, we get the particle displacements
in between at $a$ using interpolation with the cubic Hermite spline,
\begin{align}
  \vx(a) =\ &h_{00}(\alpha)\vx_i + h_{10}(\alpha)(a_{i+1} - a_i)\vv_i + \nonumber\\
           &h_{01}(\alpha)\vx_{i+1} + h_{11}(\alpha)(a_{i+1} - a_i)\vv_{i+1} \,,
\end{align}
where $\alpha := (a - a_i)/(a_{i+1} - a_i)$ and the Hermite basis functions read
\begin{align}
  h_{00}(\alpha) &= 2\alpha^3 - 3\alpha^2 + 1 \,,\nonumber\\
  h_{10}(\alpha) &= \alpha^3 - 2\alpha^2 + \alpha \,,\nonumber\\
  h_{01}(\alpha) &= -2\alpha^3 + 3\alpha^2 \,,\nonumber\\
  h_{11}(\alpha) &= \alpha^3 - \alpha^2 \,.
\end{align}
The particle velocity is then given by the derivative $\d\vx / \d a$.

\subsection{Objective Function}

Again combine Lagrangian and Eulernian fields.

A mean squared error (MSE) loss on the fields would focus too much on
the small scales. In this work we choose uniform weights in the scale
space, by weighing each mode of wavenumber $k$ by $k^{-3}$ in the MSE
loss
\begin{equation}
  \cJ = \mathbb{E} \sum_\vk k^{-3}
    \bigl| \hat f(\vk) - f(\vk) \bigr|^2.
\end{equation}
